---
title: "DATA624-HW2-Forecasting"
author: "Michael Y."
date: "2/16/2020"
subtitle: "FPP-Hyndman exercises 3.1, 3.2, 3.3 and 3.8"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 3
    keep_md: yes
    md_extensions: +grid_tables
  pdf_document:
    md_extensions: +grid_tables
    toc: yes
    toc_depth: 3
    keep_md: yes
    keep_tex: yes
urlcolor: blue
linkcolor: blue
editor_options:
  chunk_output_type: inline
header-includes: 
- \usepackage{graphicx}
- \usepackage{float}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---
---

<style>
  .main-container {
    max-width: 1300px !important;
  }
</style>

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=7,scipen=999,width=130)
```

\newpage
```{r libraries}
library(fpp2)
```

\newpage
# Homework 2 - Forecasting
Do exercises 3.1, 3.2, 3.3 and 3.8 from the online Hyndman book.
Please submit both your Rpubs link as well as attach the .rmd file with your code.

## 3.1 For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance.

### `usnetelec` : Annual US net electricity generation (billion kwh) for 1949-2003 
```{r fpp-3.1a}
### plot raw data series
autoplot(usnetelec) + 
  ggtitle(paste("usnetelec (untransformed)")) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="red")
usnetelec.ljung <- checkresiduals(naive(usnetelec))
#usnetelec.ljung

if (usnetelec.ljung$p.value > 0.05) {
  print("Because the p-value on the Ljung-Box test is large, 
        the Box-Cox transform is not necessary, but here goes:")
}  else  {
    print("Because the p-value on the Ljung-Box test is small, 
          we'll try Box-Cox transform to see if we can achieve constant variance")}

### Box-Cox transform
usnetelec.lambda <- BoxCox.lambda(usnetelec)
### Plot transformed series
#print(paste("Box-Cox lambda for usnetelec: ", round(usnetelec.lambda,3)))
autoplot(BoxCox(usnetelec, usnetelec.lambda)) + 
  ggtitle(paste("Box-Cox lambda for usnetelec: ", round(usnetelec.lambda,4))) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="blue")
  
usnetelec.xform.ljung <- checkresiduals(naive(BoxCox(usnetelec, usnetelec.lambda)))
#usnetelec.xform.ljung

```

\newpage
### `usgdp` : Quarterly US GDP. 1947:1 - 2006.1 
```{r fpp-3.1b}
### plot raw data series
autoplot(usgdp) + 
  ggtitle(paste("usgdp (untransformed)")) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="red")
usgdp.ljung <- checkresiduals(snaive(usgdp))
#usgdp.ljung

if (usgdp.ljung$p.value > 0.05) {
  print("Because the p-value on the Ljung-Box test is large, 
        the Box-Cox transform is not necessary, but here goes:")
}  else  {
    print("Because the p-value on the Ljung-Box test is small, 
          we'll try Box-Cox transform to see if we can achieve constant variance")}

### Box-Cox transform
usgdp.lambda <- BoxCox.lambda(usgdp)
### Plot transformed series
#print(paste("Box-Cox lambda for usgdp: ", round(usgdp.lambda,3)))
autoplot(BoxCox(usgdp, usgdp.lambda)) + 
  ggtitle(paste("Box-Cox lambda for usgdp: ", round(usgdp.lambda,4))) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="blue")
  
usgdp.xform.ljung <- checkresiduals(snaive(BoxCox(usgdp, usgdp.lambda)))
#usgdp.xform.ljung
```

\newpage
### `mcopper` : Monthly copper prices. Copper, grade A, electrolytic wire bars/cathodes,LME,cash (pounds/ton)  
```{r fpp-3.1c}
### plot raw data series
autoplot(mcopper) + 
  ggtitle(paste("mcopper (untransformed)")) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="red")
mcopper.ljung <- checkresiduals(naive(mcopper))
#mcopper.ljung

if (mcopper.ljung$p.value > 0.05) {
  print("Because the p-value on the Ljung-Box test is large, 
        the Box-Cox transform is not necessary, but here goes:")
}  else  {
    print("Because the p-value on the Ljung-Box test is small, 
          we'll try Box-Cox transform to see if we can achieve constant variance")}

### Box-Cox transform
mcopper.lambda <- BoxCox.lambda(mcopper)
### Plot transformed series
#print(paste("Box-Cox lambda for mcopper: ", round(mcopper.lambda,3)))
autoplot(BoxCox(mcopper, mcopper.lambda)) + 
  ggtitle(paste("Box-Cox lambda for mcopper: ", round(mcopper.lambda,4))) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="blue")
  
mcopper.xform.ljung <- checkresiduals(naive(BoxCox(mcopper, mcopper.lambda)))
#mcopper.xform.ljung
```

\newpage
### `enplanements` : Monthly US domestic enplanements  (millions): 1996-2000  
```{r fpp-3.1d}
### plot raw data series
autoplot(enplanements) + 
  ggtitle(paste("enplanements (untransformed)")) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="red")
enplanements.ljung <- checkresiduals(snaive(enplanements))
#enplanements.ljung

if (enplanements.ljung$p.value > 0.05) {
  print("Because the p-value on the Ljung-Box test is large, 
        the Box-Cox transform is not necessary, but here goes:")
}  else  {
    print("Because the p-value on the Ljung-Box test is small, 
          we'll try Box-Cox transform to see if we can achieve constant variance")}

### Box-Cox transform
enplanements.lambda <- BoxCox.lambda(enplanements)
### Plot transformed series
#print(paste("Box-Cox lambda for enplanements: ", round(enplanements.lambda,3)))
autoplot(BoxCox(enplanements, enplanements.lambda)) + 
  ggtitle(paste("Box-Cox lambda for enplanements: ", round(enplanements.lambda,4))) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="blue")
  
enplanements.xform.ljung <- checkresiduals(snaive(BoxCox(enplanements, enplanements.lambda)))
#enplanements.xform.ljung
```

***
\newpage
## 3.2 Why is a Box-Cox transformation unhelpful for the cangas data? 

### `cangas` : Monthly Canadian gas production, billions of cubic metres, January 1960 - February 2005
```{r fpp-3.2}
### plot raw data series
autoplot(cangas) + 
  ggtitle(paste("cangas (untransformed)")) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="red")
cangas.ljung <- checkresiduals(rwf(cangas,h=2*frequency(cangas),drift=T))
#cangas.ljung

if (cangas.ljung$p.value > 0.05) {
  print("Because the p-value on the Ljung-Box test is large, 
        the Box-Cox transform is not necessary, but here goes:")
}  else  {
    print("Because the p-value on the Ljung-Box test is small, 
          we'll try Box-Cox transform to see if we can achieve constant variance")}

### Box-Cox transform
cangas.lambda <- BoxCox.lambda(cangas)
### Plot transformed series
#print(paste("Box-Cox lambda for cangas: ", round(cangas.lambda,3)))
autoplot(BoxCox(cangas, cangas.lambda)) + 
  ggtitle(paste("Box-Cox lambda for cangas: ", round(cangas.lambda,4))) +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="blue")
  
cangas.xform.ljung <- checkresiduals(rwf(BoxCox(cangas, cangas.lambda),h=2*frequency(cangas),drift=T,lambda=cangas.lambda))
#cangas.xform.ljung
```


The raw data series for `cangas` exhibits much higher variance during the middle years (late 1970s through early 1990s) and lower variance in the early and later years.  The Box-Cox transformation is unable to stabilize this pattern.

***
\newpage
## 3.3. What Box-Cox transformation would you select for your retail data (from Exercise 3 in Section 2.10)?


```{r fpp3.3}
#### You can read the data into R with the following script:
#### readxl does not read straight from URL without local download
####retaildata <- readxl::read_excel("https://otexts.com/fpp2/extrafiles/retail.xlsx", skip=1)
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
#### The second argument (`skip=1`) is required because the Excel sheet has two header rows.

##myts <- ts(retaildata[,"A3349873A"],
##  frequency=12, start=c(1982,4))
#### Select one of the time series as follows 
#### (but replace the column name with your own chosen column):

mycode <- "A3349396W"
mytitle <-  "Monthly Turnover;Total(State);Total(Industry)"
mymain <- paste(mycode,mytitle)
myts <- ts(retaildata[,"A3349396W"],
  frequency=12, start=c(1982,4))


### plot raw data series
autoplot(myts) + 
  ggtitle(paste(mymain, "(untransformed)")) +
  ylab(mycode)+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="red")

myts.ljung <- checkresiduals(snaive(myts))
#myts.ljung

if (myts.ljung$p.value > 0.05) {
  print("Because the p-value on the Ljung-Box test is large, 
        the Box-Cox transform is not necessary, but here goes:")
}  else  {
    print("Because the p-value on the Ljung-Box test is small, 
          we'll try Box-Cox transform to see if we can achieve constant variance")}

### Box-Cox transform
myts.lambda <- BoxCox.lambda(myts)
### Plot transformed series
#print(paste("Box-Cox lambda for myts: ", round(myts.lambda,3)))
autoplot(BoxCox(myts, myts.lambda)) + 
  ggtitle(paste("Box-Cox lambda for", mycode, ": ", round(myts.lambda,4))) +
  ylab(paste(mycode," (transformed)"))+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_line(color="blue")
  
myts.xform.ljung <- checkresiduals(snaive(BoxCox(myts, myts.lambda)))
#myts.xform.ljung

```

The Box-Cox transformation selected for this data series is $\lambda = `r  round(myts.lambda,4)`$ .





***
\newpage
## 3.8 For your retail time series (from Exercise 3 in Section 2.10):

### a) Split the data into two parts using
```{r fpp-3.8a}
myts.train <- window(myts, end=c(2010,12))
myts.test <- window(myts, start=2011)
```



\newpage
### b) Check that your data have been split appropriately by producing the following plot.

```{r fpp-3.8b}
autoplot(myts) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")  + 
  ggtitle(mymain)+
  ylab(mycode)
```

\newpage
### c) Calculate forecasts using `snaive` applied to `myts.train`.

##### Note: To get 36 months of forecast (3 years) we have to specify h=36    
##### If we don't specify a value for h then we will get the default, which is only 2 years (24 months)     

```{r fpp-3.8c}
fc <- snaive(myts.train,h=length(myts.test))
fc
```


\newpage
### d) Compare the accuracy of your forecasts against the actual values stored in myts.test.

```{r fpp-3.8d}
accuracy(fc,myts.test)

autoplot(window(myts,start=2009)) +
  ggtitle(paste("Actual (blue and green) and prediction (red) for", mycode))+
  theme(plot.title = element_text(hjust = 0.5))+
  autolayer(window(myts.train,start=2009), series="Training") +
  autolayer(fc, series="prediction")+
  autolayer(myts.test, series="Test", )
```


\newpage
### e) Check the residuals.

```{r fpp-3.8e}
checkresiduals(fc)
```

#### Do the residuals appear to be uncorrelated and normally distributed?

No, the residuals exhibit strong autocorrelation across all lags.  Additionally the histogram shows much more density to the left of the mode, with a right tail.  The residuals are clearly biased as they do not account for the year-over-year upward trend observed in the actual data; a seasonal trend model would be more appropriate.

\blandscape
\newpage
### f) How sensitive are the accuracy measures to the training/test split?

Below I take the dataset and rerun the naive model a dozen times, each time moving one additional year's worth of data between training and test.  We start with the year 2000 being the cut-point between training and test, and advance one year at a time until reaching 2012. 

The summary of the sensitivity is as follows:   

As more data is moved into the TRAINING set, and less data is in the TEST set:

* The accuracy of **all TEST metrics improves:**
  + i.e., the **TEST** MAE, RMSE, ME, MAPE, MASE, and MPE **all** become **SMALLER**.
* **The change in accuracy of the **TRAINING** metrics is **not uniform:**
  + The accuracy of the **TRAINING** MAE, RMSE, and ME actually **WORSEN** (i.e., become **LARGER**), while 
  + the **TRAINING** MAPE and MPE initially **WORSEN**, but then eventually **IMPROVE**.    

The results are displayed below.

```{r fpp-3.8f}

firstyear = TRUE
## Loop through 13 years
for (year in 2000:2012) {
  myts.train <- window(myts, end=c(year,12))
  myts.test <- window(myts, start=year+1)
  fc <- snaive(myts.train,h=length(myts.test))
  #print(length(fc$mean))
  #print(paste("YEAR: ", year, "TRAINING SIZE: ", length(myts.train), "TEST SIZE: ",length(myts.test) ))
  ac=accuracy(fc,myts.test)
  if(firstyear == TRUE) {
    # split the "train" and "test" metrics out into two separate matrices for train and test accuracy:
    trainac  = c(YEAR=year,TRAINSIZE=length(myts.train),ac[1,])
    testac   = c(YEAR=year,TESTSIZE=length(myts.test),  ac[2,])
    firstyear = FALSE
  }
  else {
    # append the results from this year onto the existing matrices for train accuracy and test accuracy:
    trainac = rbind(trainac, c(YEAR=year,TRAINSIZE=length(myts.train),ac[1,]))
    testac  = rbind(testac,  c(YEAR=year,TESTSIZE=length(myts.test),  ac[2,]))
  }
    
}
```

\newpage
#### Accuracy metrics for TRAIN data set

```{r trainac-sensitivity-1}
# display the results of the Train Accuracy matrix
print(trainac)
# make it into a data frame, for easier plotting
traindf <- as.data.frame(trainac)
traindf$YEAR <- as.integer(traindf$YEAR)
traindf$DATE <- ISOdate(traindf$YEAR,01,01)

colors <- c("RMSE" = "blue", "MAE" = "red", "ME" = "green")

ggplot(traindf, aes(x = DATE)) +
    geom_line(aes(y = MAE, color = "MAE"), linetype="dotted", size = 1.5) +
    geom_line(aes(y = RMSE, color = "RMSE"), size = 1.5) +
    geom_line(aes(y = ME, color = "ME"), linetype="dashed", size = 1.5) +
    labs(x = "Year of Train/Test Split",
         y = "Metric",
         color = "Legend") +
    scale_color_manual(values = colors)+
  xlim(ISOdate(2000,01,01),ISOdate(2012,01,01))+
  ggtitle("Accuracy Metrics for TRAIN data set based upon date of Train/Test split")
```


```{r trainac-sensitivity-2}

colors <- c("MPE" = "blue", "MAPE" = "red", "MASE" = "green")

ggplot(traindf, aes(x = DATE)) +
    geom_line(aes(y = MPE, color = "MPE"), linetype="dotted", size = 1.5) +
    geom_line(aes(y = MAPE, color = "MAPE"), linetype="dashed", size = 1.5) +
    labs(x = "Year of Train/Test Split",
         y = "Metric",
         color = "Legend") +
    scale_color_manual(values = colors)+
  xlim(ISOdate(2000,01,01),ISOdate(2012,01,01))+
  ggtitle("Accuracy Metrics for TRAIN data set based upon date of Train/Test split")
```


\newpage
#### Accuracy metrics for TEST data set

```{r testac-sensitivity-1}
print(testac)
testdf <- as.data.frame(testac)
testdf$YEAR <- as.integer(testdf$YEAR)
testdf$DATE <- ISOdate(testdf$YEAR,01,01)


colors <- c("RMSE" = "blue", "MAE" = "red", "ME" = "green")

ggplot(testdf, aes(x = DATE)) +
    geom_line(aes(y = MAE, color = "MAE"), linetype="dotted", size = 1.5) +
    geom_line(aes(y = RMSE, color = "RMSE"), size = 1.5) +
    geom_line(aes(y = ME, color = "ME"), linetype="dashed", size = 1.5) +
    labs(x = "Year of Train/Test Split",
         y = "Metric",
         color = "Legend") +
    scale_color_manual(values = colors)+
  xlim(ISOdate(2000,01,01),ISOdate(2012,01,01))+
  ggtitle("Accuracy Metrics for TEST data set based upon date of Train/Test split")
```


```{r testac-sensitivity-2}
colors <- c("MPE" = "blue", "MAPE" = "red", "MASE" = "green")

ggplot(testdf, aes(x = DATE)) +
    geom_line(aes(y = MPE, color = "MPE"), linetype="dotted", size = 1.5) +
    geom_line(aes(y = MASE, color = "MASE"), size = 1.5) +
    geom_line(aes(y = MAPE, color = "MAPE"), linetype="dashed", size = 1.5) +
    labs(x = "Year of Train/Test Split",
         y = "Metric",
         color = "Legend") +
    scale_color_manual(values = colors)+
  xlim(ISOdate(2000,01,01),ISOdate(2012,01,01))+
  ggtitle("Accuracy Metrics for TEST data set based upon date of Train/Test split")
```

\elandscape